\documentclass[12pt]{article}
\usepackage[left=1in, right=1in, top=1in, bottom=1in]{geometry}
\usepackage{amsthm}
\usepackage{hyperref}
\usepackage{amsmath}
\setcounter{tocdepth}{1}
\usepackage{amsthm}
\usepackage[english]{babel}
\newcommand{\N}{\mathbb{N}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{claim}[theorem]{claim}
\newtheorem{prop}{Proposition}
\usepackage{tikz}
\usetikzlibrary{automata,positioning,arrows}
\setcounter{section}{1}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{enumitem}
\newtheorem*{claim*}{claim}
\usepackage{enumitem}

\begin{document}
\begin{center}
\section*{$\sim$ Assignment 5  Computational Models Spring 2022 $\sim$}
\subsubsection*{Saar Barak}
\end{center}
\subsection*{Problem 1} 
Assuming $P\ne NP$
\begin{enumerate}[label=(\alph*)]
\item Input:  Sets $A_1, ..., A_n$, and a number $k$.\\ Question: do there exist $k$ mutually disjoint sets $A_{i1}
, ..., A_{ik}$\\
\textbf{ not in  $\mathcal{P}$}
\begin{proof}
we can reduce IS $ \le_P$ MDS as follow.  let $G(V,E)$ denote an instant of IS problem, for any $v_i\in V$ generate set $A_i$ that contain all edges that have an end in $v_i$  i.e $v_iv_j \in A_i \Leftrightarrow v_iv_j  \in E $. Its immediate that there is IND-SET  $v_{i1}
, ..., v_{ik}$ size k in $G$ iff  there exist $k$ mutually disjoint sets $A_{i1}, ..., A_{ik}$. The reduction $f$ is  poly-time computable function in  $O(|V|+|E|)$, since we just need travel on $G$ and add the elements  to the sets as described above. Additionally, given a 
sets, we can verify if they are disjoint and there are at least $k$ of them in poly
time. Hence MDS$\notin P$
\end{proof}
\hrule
\item Input:  Sets $A_1, ..., A_n$.\\ Question: do there exist 3 mutually disjoint sets $A_{i}
,A_{j}, A_{k}$\\
\textbf{ in  $\mathcal{P}$}
 
\begin{claim*} $3$IND\footnote{decide if given graph have independence set size 3}$\in \mathcal{P}$ 
\end{claim*}
\begin{proof}First notice 3IND$\in NP$ using same verifier of IS.
Given $G(V,E)$ denote $|V|=n,|E|=m$ we can label its vertex $1\dots n$. run STCON on $\langle G,v_i,v_j \rangle$ for any $i,j$
.now by choose  any triplet from $\{1\dots n\}^3$ and check if  STCON$\langle G,v_i,v_j \rangle$ return 0 for any $i,j\in \{i,j,l\}$ in total we run  $n \choose 2 $ times STCON save result on the tape size   $n \choose 2 $ and check the nation for any triplet in total $n \choose 3$. all the following is can be compute in poly time. hence 3-IND$ \in $P   

\end{proof}
Hence using same reduction described in (a) its holds that $3-$IND$\le_P3-$MDS.\\
\hrule
\item Input:  Sets $A_1, ..., A_n$, and a number $k$.\\ Question: do there exist $k$  sets such that $A_{i1} 
, ..., A_{ik}$ such that $A_i\cap A_j\neq \emptyset $\\
\textbf{ not in  $\mathcal{P}$}\begin{proof}
we can reduce CLIC $ \le_P$ Q$_{(c)}$ as follow. let $G=(V,E)$  denote an instant of CLIC problem. using same proses as described at (a) we generate the sets  $A_{vi}, ..., A_{vn}$. Hence if $G$ have clic size $k$ then exists $v_1\dots  v_k$ s.t $(v_1 v_2)(v_2,v_3)  \dots (v_{k-1} v_k)\in E \Leftrightarrow$ for any $e=v_iv_j$ exists $x\in A_i,A_J$  $\Leftrightarrow$   $A_i\cap A_j$ for any $i\ne j$.
\end{proof}
\pagebreak
\item  CNF$_{50v}$\\
Input: a CNF formula $\psi$ with at most 50 variables.
\\Question: does there exist an assignment that satisfies  $\psi$ 
\\\textbf{ in  $\mathcal{P}$}
 \begin{proof}  Let  $\mathcal{A}$ denote algorithm  that get as input  CNF formula $\psi$. If $\psi$ contain more then 50 variables then $\mathcal{A}$ immediately reject, otherwise its "brute force" by set  any possible combination of boolean assignment for the given variables, If one of them satisfy $\psi$ its ACCEPT, and if none of them satisfy  $\psi$  then REJECT. Since from 50 variables  there are  total of  $2^{50}$ possible assignment, $\mathcal{A}$  is polynomial   algorithms that decide our problem.\end{proof}
\hrule 
\item Input: a CNF formula $\psi$ with at most 50 clauses.
\\Question: does there exist an assignment that satisfies  $\psi$ 
\\ \textbf{ in  $\mathcal{P}$}
\begin{proof}
Let $\psi$ be  a CNF formula with at most 50 clauses on $n$ literals. Consider some algorithm $\mathcal{A}(\psi)$ that  try all possible ways to choose one  literal from each clause whenever  $x_i,\neg x_i $ does \textbf{Not} appear together $\forall i$. If $\mathcal{A}$ find such an assignment $\Rightarrow\psi$ satisfies, otherwise it is not .Since there are at most $(2n)^{50}$ ways to choose such an assignment, $\mathcal{A}$ is polynomial   algorithms that decide our problem.
\end{proof}
\hrule 
\item 
Input: a 3CNF formula $\psi$ with even number of variables.
\\Question: does there exist an assignment that satisfies  $\psi$   and gives  $True$ 
for exactly one half of the variables?
\\ \textbf{ not in  $\mathcal{P}$}
\begin{proof}
we can reduce 3SAT $ \le_P$ 3CNF$_{half}$ as follow. Given 3SAT formula $\psi$ that have $n$ variable $x_1\dots x_n$, now construct new $n$ variable $y_1\dots y_n$ such that each $y_i$ correspond to the opposite value of  $x_i$, its can done by adding to $\psi$ 2-CNF clauses i.e \[(x_1 \vee y_1)(\overline{x_1}\vee \overline{y_1})\dots(x_n \vee y_n)(\overline{x_n}\vee \overline{y_n})\]
Any solutions to the formula will have half the variables with true values and half false. the following can done in poly-time with  computable function just duplicate  $\psi$ variables  and add the  2-CNF clauses. 
\end{proof}
\hrule 

\item 
Input: undirected graph $G$, a number $k$.
\\Question: does there exist in $G$ a clique of size at least $k$ or an independent set of size at least $k$?
\\ \textbf{ not in  $\mathcal{P}$}\begin{proof} we can reduce IS $ \le_P$ CLIC$\vee$IS as follow. let $G=(V,E)$ instant of IS problem, we can construct new graph $G'$ by adding set of additional $|V'|=|V|$ isolated vertexes. now define \[f(\langle G=(V,E) ,k \rangle)=\langle G'=(V+V',E) ,k+|V| \rangle\]
$f$ is polynomial. Since $k+|V|>|V|$ then $G'$ don't have CLIC size $k+|V|$ for sure. and exist IS size $k$ in $G$ $\Leftrightarrow$ exist IS size $k+|V|$ in $G'$. Hence CLIC$\vee$IS $\notin$P
\end{proof}
\end{enumerate}
\subsection*{Problem 2} \begin{enumerate}[label=(\alph*)]
\item \fbox{\fbox { \parbox { .55\linewidth} {For every nontrivial $L_1,L_2\in P, L_1\le L_2$ \\\textbf{TRUE}}}}

Let $L_1,L_2\in P$ W.L.O.G we can generate arbitrary $f$ that yield $L_1\le_P L_2$ 
\begin{algorithm}
\caption{ $f$ on input $w$ .}\label{alg:cap}
\begin{algorithmic} 
\State  \textbf{Decide}  $w\in L_1$
\State \textbf{If} $w\in L_1$  Return any $w'\in L_2$
\State \textbf{If} $w\notin L_1$  Return any $w''\notin L_2$
\end{algorithmic}
\end{algorithm}
\\First line can  done in poly-time since $L_1\in P$, 2$^{\text{nd}}$ holds  because $L_2\ne \emptyset$ and 3$^{\text{nd}}$ since
$L_2\ne \Sigma ^*$. All the following can done in poly-time hence $f$  poly-time computable reduction function from $L_1\le_P L_2$ \\
\item \fbox{\fbox { \parbox { .55\linewidth} {For every nontrivial $L_1,L_2\in NP, L_1\le_P L_2$ \\\textbf{Equivalent to an
Open Problem}}}}
\begin{claim*} $P=NP$ $ \Leftrightarrow$  claim $2(b)$ is True
\end{claim*}
\begin{proof}$\Rightarrow$ if $P=NP$ any $L_1,L_2\in NP \Rightarrow L_1,L_2\in P \Rightarrow $  Using 2(a)  any non-trivial $L_1,L_2 \le P$  can reduce to any other $\Rightarrow L_1\le_pL_2$ claim 2(b) Is true.
\\\\ $\Leftarrow$ Consider  some non trivial $L_2\in P$,  and assume that exist non trivial $L_1\in NP/P$. Since $P\subseteq NP$ then $L_2\in NP \Rightarrow$ claim 2(b) yield that exists reduction s.t $L_1 \le_P  L_2$ then $L_1\in P$ but  $L_1\in NP/P $ and we get an contradiction $\Rightarrow$ there is no exist such $L_1\Rightarrow NP/ P=\emptyset \Rightarrow NP\subseteq P\Rightarrow P=NP$
\end{proof}
\item \fbox{\fbox { \parbox { .55\linewidth} {$L=\{0^n1^n|n \in N \}$ is NPC \\\textbf{FALSE}}}}
\begin{claim*} $L\in\mathcal{LOGSPACE}$ 
\end{claim*}
\begin{proof}
$L$ can decided by $O(\log(n))$-space TM $\mathcal{M}$ on input $w$, that work as follow:
\begin{list}{•}{}
\item Check that 1 is ever followed by a 0
\item Count the number of 0’s and 1’s.
\item Compare the two counters.\end{list}
First step requires no working space, just moving the head.for the second we can set 2 binary counters each size $\log(n)$, and third no need extra space.
\\Hence $L\in\mathcal{LOGSPACE}\Rightarrow L \notin$ NPC\end{proof}

\item \fbox{\fbox { \parbox { .95\linewidth} { There exists a language in $ \mathcal{RE}$ that is complete w.r.t polynomial-time
reductions.\\
\textbf{TRUE}.}}}
\begin{claim*} $A_{TM}$ is complete w.r.t polynomial-time reductions.
\end{claim*}
\begin{proof} consider $f(w)=\langle \mathcal{M} ,w \rangle$ for any
 $L\in \mathcal{RE}$ and TM $\mathcal{M}$ that accept $L$. $f$ computable, and able to write the encode of $\mathcal{M} ,w $  in poly time.  Since $w\in L \Leftrightarrow 
\langle \mathcal{M} ,w \rangle \in A_{TM} $,  $f$ define poly-time reduction for any arbitrary $L \in \mathcal{RE}$. $A_{TM}$ is complete w.r.t polynomial-time reductions\end{proof}
\item \fbox{\fbox { \parbox { .95\linewidth} { If there exists a deterministic TM that decides SAT in time $n^{O(\log n)}$
\\Then every $L\in NP$ is decidable by a deterministic TM in time $n^{O(\log n)}$.
\\
\textbf{TRUE}.}}}
\begin{proof}
Let $L\in \mathcal{NP}$ and $\mathcal{M}_{SAT}$ denote D-TM  that decide SAT. Since $SAT\in\mathcal{NPC}$ then exist some poly-time reduction $f$ s.t  $L\le_PSAT$ decide SAT. Let look at d-TM $\mathcal{M}'$.
\begin{algorithm}
\caption{ $\mathcal{M}'$ on input $w$ .}\label{alg:cap}
\begin{algorithmic} 
\State  \textbf{Computes}  $f(w)$
\State \textbf{Simulate} $\mathcal{M}_{SAT}(f(w))$
and \textbf{Answer} like $\mathcal{M}_{SAT}$
\end{algorithmic}
\end{algorithm}\\
$w\in L \Leftrightarrow \mathcal{M}_{SAT}$ accept $f(w) \Leftrightarrow f(w) \in SAT$. Since $f\in \mathcal{O}(n)\Rightarrow |f(w)|=n^k$ and $\mathcal{M}_{SAT}\in \mathcal{O}{(n^{O(\log n)})}$ its following that
\[ \mathcal{M}'(w)=\mathcal{M}_{SAT}(f(w))\in\footnote{Not realy sure if its the proper way to write it.}\mathcal{O}(\mathcal{M}_{SAT}(n^k))= \mathcal{O}(n^k)^{O(\log n^k)}=\mathcal{O}(n^{O(\log n)})
\]
\end{proof}
\end{enumerate}
\pagebreak
\subsection*{Problem 3} \begin{claim*} if P = NP, there exists a polynomial-time TM, that given a
$3$CNF formula $\psi$ , outputs a satisfying assignment for $\psi$  or indicates one does
not exists.
\end{claim*}
\begin{proof}
first notice that if $P=NP$ then $NP\subseteq P$, since $3CNF\in NP$ then $3CNF\in P$. Its following that exist  polynomial-time TM $\mathcal{M}_{3CNF}$ that decide 3CNF.  Now let us define TM $\mathcal{M'}$ that given  a formula $\psi=\alpha(x_1,x_2\dots x_N)$ work such that:
\begin{algorithm}
\caption{ $\mathcal{M}'$ on input $\psi$ .}\label{alg:cap}
\begin{algorithmic} 
\State  check if   $\psi$ is valid 3CNF formula, otherwise \textbf{Reject}
\State \textbf{Simulate} $\mathcal{M}_{3CNF}\langle \psi\rangle$ and
\textbf{Reject} if $\mathcal{M}_{3CNF}\langle\psi\rangle$  \textbf{Reject}
\State $\alpha(x_1,x_2\dots x_N)\gets \psi$\Comment{$\alpha$ represent the logic relation w.r.t $\psi$}
\State $ y_j\gets 0 \quad  \quad 1\le j \le N$
\State $i\gets 1$
\While{$i \le N$}
\State \textbf{Simulate} $\mathcal{M}_{3CNF}\langle\alpha(y_1,y_2\dots y_{i-1}, T, x_{i+1}\dots  x_N)\rangle$
\If{$\mathcal{M}_{3CNF}$ Accept}\State $y_i \gets T$
\Else{}\Comment{$\mathcal{M}_{3CNF}$ reject}
\State $y_i\gets F$
\EndIf
\State $\alpha(x_i)\gets y_i$
\EndWhile
\State  \textbf{output} $y_i\dots y_N$
\end{algorithmic}
\end{algorithm}\\
If $\psi$ does not have an boolean satisfy assignment then $\mathcal{M}'$ indicate that at step 2. 
The correctness of $\mathcal{M}'$ following from the "greedy" posses, i.e before any iteration $\alpha$ can be satisfied. Hence by assign each time one of the veritable to T, we check if $\mathcal{M}_{3CNF}$ accept. if its accept the assignment its satisfied, and if its reject  then the value $F$    is the valid assignment. $\mathcal{M}'$ run $N+1 \times \mathcal{M}_{3CNF}$ that is poly-time TM that output a satisfying assignment for $\psi$  or indicates one does not exists.

\end{proof}
\subsection*{Problem 4} 
We say that a polynomial reduction $f$ is a $shrinking$ $reduction$ if there exists
$n_0$ such that for every $x\in \Sigma^*$
such that $n_0\le x , |f(x)|\le|x|-1$. Assuming $P\ne NP$
\begin{enumerate}[label=(\alph*)]
\item \fbox{\fbox { \parbox { .95\linewidth} {For every two nontrivial languages $A,B\in P$ there exists a $shrinking$ 
$reduction$ from A to B.\\\textbf{Prove}}}}
\pagebreak\begin{proof}
using Q2(a) between any non trivial $A,B\in P$ exists mapping reduction. Since both non trivial, then exists some $b \in B,\overline{b}\notin B$. Consider $f$ s.t 
\[
f(w)=\begin{cases} 
      b &\quad\text{if  } w \in A \\
        \overline{b} &\quad\text{if  } w \notin A \\ \end{cases}  
     \]
     Its immediate that $f(w)\in B \Leftrightarrow f(w)=b \Leftrightarrow w\in A  $. and $f$ define valid poly- reduction. Let us define $n_0=\max \{|\overline{b}|,|b|\} +1$, we can notice that $f$ define $shrinking$ 
$reduction$ from A to B since
\[
x\in \Sigma^* \text{ s.t }n_0\le |x| \qquad |f(x)|\le \max \{|\overline{b}|,|b|\}\le n_0-1 \le |x|-1
\]
\end{proof}
\item \fbox{\fbox { \parbox { .95\linewidth} {For every two nontrivial languages $A,B\in NPC$ there exists a $shrinking$ 
$reduction$ from A to B.\\\textbf{Disprove}}}}
\\
\begin{proof}
let $A$ be any non trivial language s.t $A\in NPC$. By consider the following claim is true, W.L.O.G we can choose $A=B$. then exist  $shrinking 
-reduction$ $f$ with some $n_0$ such that $A\le_P A$. First notice that there is only finite many $w$ s.t $|w|< n_0$. We can encode them all correct answers   to an  algorithm $\mathcal{A}$.  For given $x\in \Sigma ^*$ if $|x|< n_0$ then its encoded already to $\mathcal{A}$. Since $f(x)<|x|$, when $|x|\ge  n_0$ we can apply finite time of $ff\dots (x)$  until we achieve  some $|w|<n_0$. All can done in poly-time since $f$ is poly-reduction. Hence $\mathcal{A}$ decide any non trivial $A\in NPC$ in polynomial time, its following that $P=NP$, Contradiction. 
 \end{proof}
\end{enumerate}
\subsection*{Problem 5} 
 The following languages are NPC:
\begin{enumerate}[label=(\alph*)]
\item \[ EXACT3SAT=\{\varphi \in 3SAT:\text{every clause of $\varphi$ has exactly 3 distinct variables}\}
\]
The verifier for SAT is valid poly-time verifier for   EX-3SAT$\in NP $. We can reduce 3SAT$\le_P$EX-3SAT. Given instant of SAT $\psi=C_1\wedge C_2 \dots C_n$  for any clause $C_i$, define $f$  work as follows
\begin{list}{•}{}
\item If $C=\emptyset$ i.e $\psi$ have empty clause then it is unsatisfiable. then $f$ return any possible combination  that can generate using 3 distinct variables i.e $2^3$ clauses of 3EX-SAT
\begin{align*}
f(\emptyset)&=(x\vee y\vee z)\wedge (x\vee y\vee \neg z)\wedge (x\vee \neg y\vee z)\wedge(x\vee \neg y\vee \neg z)\wedge
\\ &(\neg x\vee y\vee z)\wedge(\neg x\vee y\vee \neg z)\wedge(\neg x\vee \neg y\vee z)\wedge(\neg x\vee \neg y\vee \neg z)
\end{align*}
$\emptyset \equiv f(\emptyset)$
\item $C=(x)$ when C have just one literal. Let us define $f$ such that 
\[f((x))=(x\vee y \vee z)\wedge 
(x\vee \neg y \vee z)\wedge 
(x\vee y \vee \neg z)\wedge 
(x\vee\neg y \vee \neg z)
\]
$(x)\equiv f(x)$
\item $C=(x\vee y)$ when C have just 2 literal. Let us define $f$ such that 
\[f((x\vee y))=(x\vee y\vee z)\wedge(x\vee y\vee \neg z)
\]
$(x\vee y)\equiv f((x\vee y))$
\item $C=(x\vee y \vee z)$ then $f(C)=C$
\end{list}
Hence let  $f'(\psi)=f'(C_1\wedge C_2 \dots C_n)=f(C_1)\wedge f(C_2)\dots f(C_n)$ define poly-time reduction from 3SAT to EX-3SAT. Following that EX-3SAT$\in NPC$ 
\item \[ L_2=\{\langle M,1^n \rangle :\text{M is a TM and there exists a string that M accepts in n steps}\}
\]
\begin{proof} 
First consider some $w=\langle M,1^n \rangle $ we can verify that $w\in L_2$, by simulate M on any input size $n$ that is\footnote{ I assumed a finite alphabet } $n^{|\Sigma ^*|}\in \mathcal{O}(n)$, and in total $\mathcal{O}(n^{|\Sigma ^*|} |M|^3n \log n ) \Rightarrow L_2\in NP$\\
Let $L$ be any $L\in NP$, and  $V_L$ denote its polynomial  verifier, and assume its runtime is $p(|x|)$. Any such $L$ can reduce $L\le_P L_2$ as follows
\[x\in L \Leftrightarrow \exists c \quad \text{s.t } (x,c)\in V_L \Leftrightarrow f(x)=\langle V_L,1^{p(|x|)}\rangle\in L_2 
\]
The correctness followed by the verifier definition, and $f$ define computable poly-reduction from any $L\in NP$, Hence $L_2\in NPC$
\end{proof} 
\end{enumerate}
\end{document}
